---
title: Origin-destination data to route networks workflows
format: gfm
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, cache = TRUE)
library(tidyverse)
library(tmap)
```

The aim of this document is to demonstrate how to convert Origin-Destination data to route networks.
The data for this is from the [wicid](https://wicid.ukdataservice.ac.uk/flowdata/cider/wicid/downloads.php) website.

You can download datasets at multiple geographic levels from that website.
We will start with the open MSOA to MSOA dataset.

The study location can be defined as follows:


```{r}
study_location_name = "Alder Hay Hospital"
study_location_coodinates = stplanr::geo_code(study_location_name)
study_location_coodinates
study_location_df = data.frame(name = study_location_name, lon = study_location_coodinates[1], lat = study_location_coodinates[2])
study_location_sf = sf::st_as_sf(study_location_df, coords = c("lon", "lat"), crs = 4326)
buffer_distance = 5000
study_location_buffer = sf::st_buffer(study_location_sf, buffer_distance)
```

For the MSOA data we will get data from the `pct` R package as follows:


```{r}
msoa_data = pct::get_pct(layer = "z", geography = "msoa", national = TRUE)
```
